---
title: "Model to predict weekly NFL Attendance"
author: "Karani K Wachira"
format: html
editor: visual
---

Load the NLF attendance data sets to use

```{r}

library(tidyverse)

attendance <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-04/attendance.csv")

standings <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-04/standings.csv")

attendance_joined <- attendance %>% 
  left_join(standings, 
            by = c("year", "team_name", "team"))

View(attendance_joined)
```

How does weekly attendance vary for different teams, and for the seasons they did/did not make the playoffs?

```{r}

attendance_joined %>% 
  filter(!is.na(weekly_attendance)) %>% 
  ggplot(aes(fct_reorder(team_name, weekly_attendance),
             weekly_attendance,
             fill = playoffs)) +
  geom_boxplot(outlier.alpha = 0.4) +
  coord_flip() +
  labs(fill = NULL, x = NULL,
       y = "weekly NLF game attendance")


```

32 teams in the NFL, there are years they all did & didn't make the playoffs.

How does margin_of_victory a measure of points scored relative to points allowed, measure the same thing as getting to the playoffs?

```{r}

attendance_joined %>% 
  distinct(team_name, year, margin_of_victory, playoffs) %>% 
  ggplot(aes(margin_of_victory,  fill= playoffs)) +
  geom_histogram(position = "identity", alpha = 0.7) +
  labs(
    x = "Margin of victory",
    y = "Number of teams",
    fill = NULL
  )



```

```{r}
attendance_joined %>% 
  mutate(week = factor(week)) %>% 
  ggplot(aes(week, weekly_attendance, fill = week)) +
  geom_boxplot(show.legend = FALSE, outlier.alpha = 0.5) +
  coord_flip() +
  labs (
    x = "week of NLF season",
    y = "week NFL game attendance"
  )
```

remove the weeks that each did not play(where weekly is NA)

```{r}

attendance_df <- attendance_joined %>% 
  filter(!is.na(weekly_attendance)) %>% 
  select(weekly_attendance, team_name, year, week, margin_of_victory, strength_of_schedule, playoffs)

attendance_df




```

Build simple models

First step is data into training and test.

Each data set will have about the same number of examples of teams that went on to the playoffs

```{r}
library(tidymodels)

set.seed(1234) 
attendance_split <- attendance_df %>% 
  initial_split(strata = playoffs)


nfl_train <- training(attendance_split)

nfl_test <- testing(attendance_split)

```

```{r}
lm_spec <- linear_reg() %>% 
  set_engine(engine = "lm")



lm_spec

```

model 1

```{r}
lm_fit <- lm_spec %>% 
  fit(weekly_attendance ~., 
      data = nfl_train)


lm_fit

```

```{r}

rf_spec <- rand_forest(mode = "regression") %>% 
  set_engine("ranger")


rf_spec
```

```{r}
library(ranger)

rf_fit <- rf_spec %>%
  fit(weekly_attendance ~ ., data = nfl_train)



rf_fit
```

Evaluate models

```{r}

results_train <- lm_fit %>% 
  predict(new_data = nfl_train) %>% 
  mutate(
    truth = nfl_train$weekly_attendance,
    model = "lm"
  ) %>% 
  bind_rows(rf_fit %>% 
              predict(new_data = nfl_train) %>% 
              mutate(
                truth = nfl_train$weekly_attendance,
                model = "rf"
              ))






```
